---
title       : Beta Distribution Explorer
author      : RaphT
framework   : revealjs
highlighter : highlight.js
hitheme     : default 
widgets     : [mathjax]
mode        : standalone
knit        : slidify::knit2slides
revealjs:
  theme: beige
---

### Beta Distribution Explorer

By RaphT, Feb 2015
<br><br>
<br><br>
<br><br>
Try out the explorer [here](http://rapht.shinyapps.io/DataProductProject/)

--- &twocol

### The Beta Distribution Family

*** =left
The Beta distribution is a family of distributions:
- Continuously defined in the [0,1] interval
- Controlled by two positive shape parameters alpha and beta
- Used to model random variables limited to a finite length interval

*** =right
```{r, echo = F, dev.args = list(bg = "transparent"),message=F, warning = F,fig.width=5, fig.height=5}
library(ggplot2)
library(betareg)
alpha = 0.5
beta = 0.5
dist = data.frame(x = rbeta(10000, alpha, beta), z = rbeta(10000, alpha*20, beta*20), w = rbeta(10000, alpha*4, beta*20))
p = ggplot(dist, aes(x=x)) + 
  geom_histogram(aes(y=..density..),
                 breaks=seq(0,1,by=0.011), 
                 colour="black", 
                 fill="white")+
  geom_histogram(aes(x = z, y=..density..),
                 breaks=seq(0,1,by=0.011), 
                 colour="red", alpha = 0.3,
                 fill="white")+
  geom_histogram(aes(x = w, y=..density..),
                 breaks=seq(0,1,by=0.011), 
                 colour="green", alpha = 0.3,
                 fill="white")+
  scale_x_continuous("")+
  scale_y_continuous( limits = c(0,7))
print(p)
```

<span style="font-size:0.5em;">Examples of beta distribution</span>

*** =fullwidth
[Beta distributions on Wikipedia](http://en.wikipedia.org/wiki/Beta_distribution)

--- 

### Explorer work flow

1. Simulates a beta distribution based on a set of parameters 
2. Uses **[Betareg](http://cran.r-project.org/web/packages/betareg/vignettes/betareg.pdf)** to fit alpha and beta on the simulated distribution 
3. Displays the distribution and the theoretical and fitted density functions 

```{r, dev.args = list(bg = "transparent"),message=F, warning = F,fig.width=5, fig.height=5}
alpha = 2; beta = 3; dist = data.frame(x = rbeta(100, alpha, beta))
```
```{r, dev.args = list(bg = "transparent"),message=F, warning = F,fig.width=5, fig.height=5}
model = betareg(x~., link = "logit", link.phi = "log", data = dist)
```
```{r, echo = F, dev.args = list(bg = "transparent"),message=F, warning = F,fig.width=3, fig.height=3}
mu.fitted = plogis(coef(model)[1])
phi.fitted = exp(coef(model)[2])
alpha.fitted = mu.fitted*phi.fitted
beta.fitted = phi.fitted*(1-mu.fitted)
```
```{r, dev.args = list(bg = "transparent"),message=F, warning = F,fig.width=5, fig.height=3}
ggplot(dist, aes(x=x)) + geom_histogram(aes(y=..density..), breaks=seq(0,1,by=0.011))+
  stat_function(fun=dbeta, args=list(shape1=alpha, shape2 = beta),col = "red")+
  stat_function(fun=dbeta, args=list(shape1=alpha.fitted, shape2 = beta.fitted),col = "green")
```

--- &twocol

### Explorer Options

*** =left2

Interactively change the value of alpha and beta via sliders
<br><br>
Interactively change the number of samples in the distribution
<br><br>
Choose the standard or alternative parametrization of the distribution

&nbsp;&nbsp;&nbsp;1. The standard directly uses alpha and beta

&nbsp;&nbsp;&nbsp;2. The alternative uses mu (location) and phi (dispersion)

--- 

### Future work

Add the possibility to open and fit custom data set
<br><br>
<br><br>
<br><br>

### Acknowledgments

This [webpage](http://zevross.com/blog/2014/11/19/creating-elegant-html-presentations-that-feature-r-code/) helped me get the two columns to work

